# Introduction to Exploratory Data Analysis & Descriptive Statistics."


**Objectives: To learn how to analyze data sets, applying methods of descriptive statistics and visualization in R.**
<br/>

"In statistics, **Exploratory Data Analysis (EDA)** is an approach to analyzing data sets to summarize their main characteristics, often with visual methods. A statistical model can be used or not, but primarily EDA is for seeing what the data can tell us beyond the formal modeling or hypothesis testing task."" (Wikipedia)

<br/>
"**Descriptive statistics** aims to summarize __a sample__, rather than use the data to learn about **the population** that the sample of data is thought to represent. This generally means that **descriptive statistics**, unlike **inferential statistics**, are not developed on the basis of probability theory."" (Wikipedia)
<br/><br/>


## Exploratory Data Analysis

We will use the package "car". For details on this package, see https://cran.r-project.org/web/packages/car/car.pdf

```{r, results="hide", warning = F}
# install.packages("car") - run this code if you do not have the "car" package installed
library(car) 
```

<br/>
Let's explore the dataset "Davis" from the car package. It is called "Self-Reports of Height and Weight Description". The subjects were men and women engaged in regular exercise. <br/>This data frame contains the following columns:
<br/>- sex    - F, female; M, male 
<br/>- weight - measured weight in kg
<br/>- height - measured height in cm
<br/>- repwt  - reported weight in kg
<br/>- repht  - reported height in cm

```{r, results="hide"}
?Davis
data <- Davis
```
<br/>

### Data dimentionality: functions _str(), summary(), head(), tail()_
```{r, results="hide"}
dim(data) 
str(data) 

head(data)
tail(data)

summary(data)  # shows quantiles for each column and how many NA !!!!
```
<br/>

### Missing (NA) values in data: functions _complete.cases(), na.omit(), all.equal()_
How many rows do not contain missing values (i.e., not a single 'NA')?
```{r, results="hide"}
sum(complete.cases(data)) 
x <- data[complete.cases(data), ] # here they are
y <- na.omit(data)
all.equal(x,y)
```

**Excercise using _complete_cases()_: How many rows contain missing values (i.e., at least one 'NA')?**
```{r, results="hide", echo=FALSE}
sum(!complete.cases(data)) 
data[!complete.cases(data), ] # here they are
```
<br/>

### Looking at the subset of data
```{r, results="hide"}
d <- data[data$weight < 60,] # rows with weight below 60 kg
str(d)
summary(d)

x <- data[data$weight > 50 & data$repwt <= 66,] 
x  # the result looks strange!?
x[!complete.cases(x),]
na.omit(x)  # we cannot do this because this removes rows that contain at least one NA in any column

y <- data[data$weight > 50 & data$repwt <= 66 & !is.na(data$repwt), ] 
y
dim(y)
y[!complete.cases(y), ]
```
<br/>

### Excercises on data subsetting and missing values

1. How many people shorter than 170 cm reported that they are taller?
```{r, results="hide"}

x <- data[data$height < 170 & data$repht >= 170 & !is.na(data$repht), ]
nrow(x)
```

2. What proportion of men in the dataset did not report their height? And women?
```{r, results="hide"}

x <- data[data$sex == 'M' & is.na(data$repht), ]
nrow(x)
nrow(x)/nrow(data[data$sex == 'M', ])

nrow(data[data$sex == 'F' & is.na(data$repht), ]) / nrow(data[data$sex == 'F', ])
```

3. Is it true that the same men who did not report height did not also report weight?
```{r, results="hide"}

all.equal(data[data$sex == 'M' & is.na(data$repht), ], data[data$sex == 'M' & is.na(data$repwt), ])
```
<br/>



### Exploring a particular variable (column): functions _unique(), table()_
```{r, results="hide"}
length(data$repwt)
summary(data$repwt)
sum(is.na(data$repwt)) # count NA
table(is.na(data$repwt))  # table of how many complete data (FALSE) and how many missing (TRUE) for data$repwt

unique(data$repwt) # shows unique values for a specified column (NA is considered)
length(unique(data$repwt))

unique(na.omit(data$repwt)) # shows unique values for a specified column (NA is omitted)
length(unique(na.omit(data$repwt)))

table(data$repwt) # how many data with the same repwt
data.frame(table(data$repwt)) # same as above shown as a data frame     

table(data$repwt, useNA = "ifany") # by default table() doesn't show missing values
data.frame(table(data$repwt, useNA = "ifany"))
```
<br/>

### Exploring relationships between variables: functions _table(), cut()_, and functions for factors _levels(), nlevels()_
```{r, results="hide"}
table(data$sex) # table() builds a contingency table of the counts at each combination of factor levels
table(data$sex, data$weight)  # shows relationships between variables
table(data$sex, data$weight < 80) # gives the 2x2 contingency table

m <- table(data$sex, data$weight < 80) # save it as a matrix
colnames(m) <- c("weight >= 80","weight < 80") # add names to columns
rownames(m) <- c("Female","Male") # add names to rows
m # check it out

table(data$weight)
intervals_weight <- cut(data$weight, breaks = seq(30, 170, 20)) 
table(intervals_weight)
table(intervals_weight, data$sex) # contigency table of sex by intervals 

class(intervals_weight)
levels(intervals_weight)
nlevels(intervals_weight)

table(data$height)
intervals_height <- cut(data$height, breaks = seq(55, 200, 20))
table(intervals_weight, intervals_height)
```
<br/>

### Excercises using _unique(), table()_ and _cut()_

Let's assume that a person with the minimum height, or _== min(data$height)_, is a wrong entry in the dataset and exclude it from the analysis. <br/>

1. How many unique values are there for the height?
```{r, results="hide"}

x <- data[!data$height == min(data$height), ]

unique(x$height)
length(unique(x$height))
```

2. How many intervals for the height will be obtained at breaks of 10 cm from minimum to maximum height. Use min() and max() in function seq() and nlevels() -- be careful to include maximum value for height in the last interval.

```{r, results="hide"}

min(x$height)
max(x$height)
intervals_height <- cut(x$height, breaks = seq(min(x$height), max(x$height)+1, 10))
nlevels(intervals_height)
```

3. How many women are in the last two intervals? (just by looking at the table)

```{r, results="hide"}

table(intervals_height, x$sex)
```
<br/>

<br/>

## Descriptive Statistics 

<br/>

### Functions: _mean, sd, var, min, max, median, range, IQR, quantile_. 
**All these function require special treatment for missing values, using parameter _na.rm = TRUE_.**

<br/>

```{r, results="hide"}
min(data$repwt) 
?min
min(data$repwt, na.rm = TRUE) 

max(data$repwt, na.rm = TRUE)
range(data$repwt, na.rm = TRUE)
range(data$repwt, na.rm = TRUE)[1] == min(data$repwt, na.rm = TRUE)
range(data$repwt, na.rm = TRUE)[2] == max(data$repwt, na.rm = TRUE)

mean(data$repwt, na.rm = TRUE) 
median(data$repwt, na.rm = TRUE)
quantile(data$repwt, na.rm = TRUE)  # shows quantiles for values in a specific column, ignoring 'NA'
quantile(data$repwt, na.rm = TRUE, probs = c(0, 0.25, 0.5, 0.75, 1)) # standard quantiles (or quartiles)
quantile(data$repwt, na.rm = TRUE, probs = c(0, 0.1, 0.9, 1)) # any other quantiles

IQR(data$repwt, na.rm = TRUE) 
IQR(data$repwt, na.rm = TRUE) == quantile(data$repwt, na.rm = TRUE)[4] - quantile(data$repwt, na.rm = TRUE)[2]

# contigency table of quantiles of weight versus quantiles of repwt
x <- data$weight
any(is.na(x))
y <- data$repwt
any(is.na(y))
table(cut(x, quantile(x)), cut(y, quantile(y, na.rm = TRUE))) 

```

<br/>

### Functions _tapply() and round()_
```{r, results="hide"}
var(data$repwt, na.rm = TRUE)
?var

sd(data$repwt, na.rm = TRUE)
sd(data$repwt, na.rm = TRUE) == sqrt(var(data$repwt, na.rm = TRUE))
sd(data$repwt, na.rm = TRUE) ** 2 == var(data$repwt, na.rm = TRUE)

# function tapply()
?tapply
tapply(data$weight, data$sex, mean)
tapply(data$weight, data$sex, quantile)
tapply(data$weight, data$sex, quantile)$F

```

<br/>

### Excercises

1. How many women have weight below (or equal) median AND height above median?
```{r, results="hide"}

d <- data[data$sex == "F", ]

# strightforward solution
nrow(d[d$weight <= median(d$weight) & d$height > median(d$height), ])

# solution using table, cut and quantile
x <- d$weight
y <- d$height
table(cut(x, quantile(x)), cut(y, quantile(y, na.rm = TRUE))) 

z <- table(cut(x, quantile(x, probs = c(0, 0.5, 1))), cut(y, quantile(y, probs = c(0, 0.5, 1))))
z[3]

```

2. How many women have weight in the lower 5% quantile AND height in the upper 5% quantile?
```{r, results="hide", echo=F}

z <- table(cut(x, quantile(x, probs = c(0, 0.05, 1))), cut(y, quantile(y, probs = c(0, 0.05, 1))))
z[3]
```

3. What is the difference of the mean and sd values of the reported height between men and women (round reported values to one digit)?
```{r, results="hide", echo=F}
x <- data[!is.na(data$repht), ]  

z <- tapply(x$repht, x$sex, mean)
z[2] - z[1]
diff <- round(z[2] - z[1], 1)
print(paste("Mean difference of reported height for men and women is", diff, "cm", sep=" "))

z <- tapply(x$repht, x$sex, sd)
diff <- z[2] - z[1]
diff
print(paste("SD difference of reported height for men and women is", round(diff, 1), "cm", sep=" "))

```

<br/>


### Data visualization: _boxplot()_ 
```{r, results="hide", fig.show="hide"}

boxplot(data$weight) 

y <- median(data$weight)
segments(0, y, 1, y, lwd = 3, col = "red")

y <- mean(data$weight)
segments(0, y, 1, y, lwd = 3, col = "blue")

y <- mean(data$weight) - sd(data$weight)
segments(0, y, 1, y, lwd = 3, col = "green")

y <- mean(data$weight) + sd(data$weight)
segments(0, y, 1, y, lwd = 3, col = "green")

y <- quantile(data$weight)[1]
segments(0, y, 1, y, lwd = 3, col = "orange")

y <- quantile(data$weight)[2]
segments(0, y, 1, y, lwd = 3, col = "magenta")

y <- quantile(data$weight)[4]
segments(0, y, 1, y, lwd = 3, col = "cyan")

y <- quantile(data$weight)[5]
segments(0, y, 1, y, lwd = 3, col = "orange")
```

<br/>

### Outliers
**Can be defined (by the Tuley's test) as values in the sample that differ from the Q1 (25% quantile) or Q3 (75% quantile) by more than 1.5 x IQR (where IQR = Q3 - Q1).** However, there is no formal definition of outliers; they need to be treated subjectively.

```{r, results="hide", fig.show="hide"}
quantile(data$weight)

lower_limit <- quantile(data$weight)[2] - 1.5 * IQR(data$weight) # values below this limit are outliers
lower_limit
min(data$weight) < lower_limit # Is the minimum weight an outlier? 

lower_whisker <- max(min(data$weight), lower_limit)
lower_whisker

boxplot(data$weight)
y <- lower_whisker
segments(0, y, 1, y, lwd = 3, col = "black")

upper_limit <- quantile(data$weight)[4] + 1.5 * IQR(data$weight) # values above this limit are outliers
upper_limit
max(data$weight) > upper_limit # Is the maximum weight an outlier? TRUE or FALSE?

upper_whisker = min(max(data$weight), upper_limit) 
upper_whisker

y <- upper_whisker
segments(0, y, 1, y, lwd = 3, col = "black")

#show outliers
data[data$weight > upper_whisker, ]
```

<br/>


### How statistics change if to remove outliers 
```{r, results="hide", fig.show="hide"}
# We can remove all "formal" outliers
d <- data[data$weight <= upper_whisker, ] 

# While it is better to remove only those that have no sense
d <- data[data$weight != max(data$weight), ] 

# Did mean change?
mean(d$weight)
mean(data$weight)

# Did median change?
median(data$weight)
median(d$weight)

# Did SD change?
sd(data$weight)
sd(d$weight)

boxplot(d$weight)
boxplot(d$weight, outline = FALSE) # boxplot has a parameter "outline" not to show outliers! 
```

<br/>

## More basic plots 
<br/>

### How to plot box-plots side-by-side on one graph
```{r, results="hide", fig.show="hide"}
boxplot(data$weight ~ data$sex)
boxplot(data$weight ~ data$sex, outline = FALSE) 
colors <- c("blue", "red")
ylab <- "weight, kg"
boxplot(data$weight ~ data$sex, outline = FALSE, border = colors, ylab = ylab)
```

<br/>

### Excercise: Make a similar boxplot using _ggplot()_
```{r, results="hide", warning = FALSE, fig.show="hide"}
#install.packages("ggplot2")
library(ggplot2) 

ggplot(data=data, aes(x=sex, y=weight)) + geom_boxplot()

p <- ggplot(data=data, aes(x=sex, y=weight, col=sex)) + geom_boxplot(outlier.shape = NA) 
p <- p + scale_y_continuous(limits = quantile(data$weight, c(0.1, 0.9)))
p <- p + theme_bw()
p <- p + ylab(label=ylab) + xlab(label="")
p <- p + scale_color_manual(values=colors)
p <- p + theme(legend.position="none")
p
```

<br/>

### How to plot together data from two or more vectors of different lengths
```{r, results="hide", warning = FALSE}
#install.packages("reshape2")
library(reshape2) # for melt() to use below for transforming a data frame from the wide to the long format
```

<br/>
```{r, results="hide", warning = F, message=F, fig.show="hide"}
# let's make two vectors first
d <- data[data$weight != max(data$weight), ]
male <- d[d$sex =="M", ] 
female <- d[d$sex =="F", ] 
m <- male$weight
f <- female$weight
length(m)
length(f)

# make a data frame on a long format
x <- data.frame( value = m, variable = rep("Male", length(m)) )
y <- data.frame( value = f, variable = rep("Female", length(f)) )
df <- rbind(x,y)

boxplot(data = df, value ~ variable)
```

<br/>

### How to make a box-plot more informative and customized
```{r, results="hide", fig.show="hide"}
# Define the plot parameters
y_limits <- c(30, 140)
colors <- c("blue", "red")
ylab <- "Weight, kg"
title <- "Distribution of weight by sex"

boxplot(data = df, value ~ variable, outline = FALSE,
        #pars = list(boxwex = .4),
        ylab = ylab,
        cex.lab = 1.5, #to change (multiply) the font size of the axes legends
        cex.axis = 1.2, #to change (multiply) the font size of the axes
        ylim = y_limits,
        border = colors, #color the boxplot borders
        boxwex = 0.6,
        staplewex = 0.4,
        frame.plot = FALSE, #this removes upper and right borders on the plot area
        outwex = 0.5,
        cex.main = 1.5, #to change (multiply) the size of the title
        main = title # the title of the graph
        )

stripchart(data = df, value ~ variable,
           col = colors,
           method = "jitter", jitter = .2,
           pch = c(16, 15), cex = c(1.0, 1.0), #different points and of different size can be used
           vertical = TRUE, add = TRUE)

# let's show the (yet unknown) p-value on the plot
text <- "p-value = ..."

y <- 130 # y position of the horizontal line
offset <- 5 # length of vertical segments
x <- 1

segments(x, y, x + 1, y)
segments(x, y - offset, x, y)
segments(x + 1, y - offset, x + 1, y)
text(x + 0.5, y + offset, paste(text), cex = 1) #cex defines the font size of the text

```
<br/>

### Histogram
```{r, results="hide", fig.show="hide"}
hist(data$weight)
hist(d$weight)
```
<br/>

### Control for the size of bins
```{r, results="hide", fig.show="hide"}
bin_size <- 5
start <- 30
end <- 120
bins <- seq(start, end, by = bin_size)
hist(d$weight, breaks = bins, col = "blue")
```

<br/>

### Two overlaying histograms on one graph

```{r, results = "hide", fig.show="hide"}
male <- subset(d, sex =="M") # we will use data with outliers removed
female <- subset(d, sex =="F")
m <- male$weight
f <- female$weight

bin_size <- 5
start <- min(min(m), min(f)) - 3*bin_size
end <- max(max(m), max(f)) + 3*bin_size
bins <- seq(start, end, by = bin_size)

xlim = c(start, end)
colors <- c(rgb(1, 0, 1, 0.7), rgb(0, 0, 1, 0.5)) # the last number in rgb() is for transparency

hist(female$weight, breaks = bins, col = colors[1], xlim = xlim, xlab = "Weight, kg")
hist(male$weight, breaks = bins, add = TRUE, col = colors[2], xlim = xlim)
legend("topright", legend = c("Females", "Males"), fill = colors, bty = "n", border = NA)
```


<br/>

### Scatterplot
```{r, results="hide", fig.show="hide"}
plot(d$weight, d$repwt, pch = 20) # Just a simple plot of reported weights against measured weights

plot(d$weight, d$repwt, pch = 20, col = d$sex) # it can be colored by sex, for example
legend("topleft",legend = c("Females", "Males"), col = 1:2, pch = 20)
```

<br/>

### Function _palette()_
Colors are taken in the order from the currently setup **palette()** . 
```{r, results="hide"}
palette()
```

<br/>


Palette's colors can be changed and then reset back to default, using palette("default").
```{r, results="hide", fig.show="hide"}
palette(c(rgb(1, 0, 1, 0.7), "blue")) # changing palette()

plot(d$weight, d$repwt, col = d$sex, pch = 20, xlab = "Measured weight, kg", ylab = "Reported weight, kg", main = "Reported versus measured weights by sex") 
legend("bottomright",legend = c("Females", "Males"), col = 1:2, pch = 20, bty = "n")

palette("default") # reset back to the default
```
<br/>


## Wrapping up everything you have learned in this course

**The goal is to read data from the file, clean and explore data, and report in the files cleaned data and the results of descriptive statistics (_number of observations, mean and SD_) of measured variables affected by two drugs independently in women and men.**

<br/>

### Read data from the file
```{r, results="hide", fig.show="hide", eval=FALSE}

system(command="svn export https://github.com/biocorecrg/CRG_RIntroduction/trunk/i_o_files")

conn <- file("./i_o_files/example_data.txt", "r")
DATA <- as.data.frame(read.table(conn, header = TRUE, sep = "\t"))
data <- DATA  # we will save DATA to return to it if needed

head(data)
tail(data)

str(data) 
dim(data)

```

```{r, results="hide", fig.show="hide", echo=FALSE, eval=F}

conn <- file("./i_o_files/example_data.txt", "r")
DATA <- as.data.frame(read.table(conn, header = TRUE, sep = "\t"))
data <- DATA  # we will save DATA to return to it if needed

```

```{r, echo=F, eval=T}
# August 2019 - added but not shown, for a proper reading of the file by bookdown
system(command="svn export --force https://github.com/biocorecrg/CRG_RIntroduction/trunk/i_o_files")
data <- as.data.frame(read.table("./i_o_files/example_data.txt", header = TRUE, sep = "\t"))
```
<br/>

### Remove empty columns 
```{r, results="hide", fig.show="hide"}

# one of the ways to remove unwanted columns
x <- data
x <- x[, grep("X.[2-9]", colnames(x), inv = T)] # be careful here because we will need columns "X" and "X.1"
x <- x[, grep("X.1[0-9]", colnames(x), inv = T)]
colnames(x)

# more elegant (and correct) way that doesn't use grep and doesn't require knowledge of columns
apply(is.na(data), 2, all)

data <- data[, !apply(is.na(data), 2, all)]

colnames(data)

str(data)
```
<br/>

### Explore, rename and clean variables
```{r, results="hide", fig.show="hide"}

# rename the colum for sex
colnames(data)[2] <- "SEX"

table(data$SEX)
# and remove rows with SEX other than M and F
data <- data[data$SEX %in% c("F", "M"), ]

# How many rows were removed?

# what is the data type of data$SEX?

class(data$SEX)
levels(data$SEX)

# remove unused levels
data$SEX <- droplevels(data$SEX)
str(data$SEX)
```
<br/>

Do the same for the column containing information on drugs
```{r, results="hide", fig.show="hide"}

colnames(data)
colnames(data)[colnames(data) == "X.1"] <- "DRUG"

data$DRUG
table(data$DRUG)

data <- data[data$DRUG %in% c("ART", "PRG"), ]
table(data$DRUG)
levels(data$DRUG)
data$DRUG <- droplevels(data$DRUG)
levels(data$DRUG)
```
<br/>


### Correct non-numeric values, changing them to NA
```{r, results="hide", fig.show="hide", warning=FALSE}

# you can correct columns one by one
x <- as.numeric(as.character(data$U_12))
str(x)
# etc....

# more elegant way to correct multiple columns at once
df <- data[, !colnames(data) %in% c("SEX", "DRUG")]

apply(df, 2, as.numeric)

x <- data.frame(apply(df, 2, as.numeric))
# well, we lost columns SEX and DRUGS, let's add them

df <- cbind(SEX=data$SEX, DRUG=data$DRUG, x)

data <- df # here is our clean dataset
str(data)

```
<br/>


### Write corrected data frame in the file
```{r, results="hide"}

conn <- file("corrected_data.txt", "w")
write.table(data, conn, row.names = F, sep = "\t", quote = F)
close(conn)

# now let's read corrected data
conn <- file("corrected_data.txt", "r")
DATA <- as.data.frame(read.table(conn, header = TRUE, sep = "\t"))
close(conn)

all.equal(DATA, data)
data <- DATA  # we will save DATA to return to it if needed

```
<br/>


### Explore and remove outliers 
We are interested in 4 groups (by sex and drug) independently
```{r, results="hide", fig.show="hide", warning=F}

summary(data)

#install.packages("ggplot2")
library(ggplot2) 

for (num in 4:ncol(data)){ # skip columns SEX, DRUG, ID
  print(colnames(data)[num])
  p <- ggplot(data=data, aes(x = SEX, y = data[ ,num], col = DRUG))  
  p <- p + geom_boxplot() + ggtitle(paste(colnames(data)[num]))
  print(p)
}

# Let's remove "obvious"" outliers by changing their values to NA

# data point below -20 and above 5 in A
data$A[data$A < -20] <- NA
data$A[data$A > 5] <- NA

# data point below -10 in I_6
data$I_6[data$I_6 <= -10] <- NA 

# data point <-4 in S_6
data$S_6[data$S_6 < -4] <- NA

# data point <-9 in I_2
data$I_2[data$I_2 < -9] <- NA

# data point above 4 in S_2
data$S_2[data$S_2 > 4] <- NA

# data point above 10 in U_2
data$U_2[data$U_2 > 10] <- NA

# data point below 10 in D1
data$D1[data$D1 <= 10] <- NA


```
<br/>


### How to change a value of a specific data point
```{r, results="hide", warning=F}
data[data$SEX=="M" & data$DRUG=="ART",]$E

# difficult way
df[df$SEX=="M" & df$DRUG=="ART", ]$E[df[df$SEX=="M" & df$DRUG=="ART", ]$E == 911] <- 500 # or NA
df[df$SEX=="M" & df$DRUG=="ART",]$E

# simpler way
df <- data
df[df$SEX=="M" & df$DRUG=="ART",]$E
df[df$SEX=="M" & df$DRUG=="ART" & df$E == 911 & !is.na(df$E), ]$E <- 500
df[df$SEX=="M" & df$DRUG=="ART",]$E
```
<br/>


### Make a data frame with statistical data
```{r, results="hide", warning=F}

#let's make first a data frame for only one group
d <- data[data$SEX == "F" & data$DRUG == "ART",]

result <- list()
names <- c("N", "mean", "sd")
count <- 1
for (i in colnames(d)){
  print(i)
  if (i == "SEX" | i == "DRUG" | i == "ID") next;
  x <- d[ ,i]
  x <- x[!is.na(x)]
  v <- c(i, length(x), mean(x), sd(x))
  result[[count]] <- v
  count <- count + 1
}

res <- data.frame(matrix(unlist(result), nrow=count-1, byrow=T))
sex <- "F"
drug <- "ART"
colnames(res) <- c("variable", paste(sex, drug, names,sep="_"))

# make a data frame for statistics for all groups, re-using the above for-loop
df <- data.frame()
names <- c("N", "mean", "sd")
for (sex in levels(data$SEX)){ 
  for(drug in levels(data$DRUG)){
    d <- data[data$SEX == sex & data$DRUG == drug,]
    result <- list()
    count <- 1
    for (i in colnames(d)){
      print(i)
      if (i == "SEX" | i == "DRUG" | i == "ID") next;
      x <- d[ ,i]
      x <- x[!is.na(x)]
      v <- c(i, length(x), mean(x), sd(x))
      result[[count]] <- v
      count <- count + 1
    }
    res <- data.frame(matrix(unlist(result), nrow=count-1, byrow=T), stringsAsFactors=FALSE)
    colnames(res) <- c("variable", paste(sex, drug, names,sep="_"))
    
    if (dim(df)[1] == 0) {
      df <- res
    }else{
      df <- cbind.data.frame(df, res)
    }
  }
}

# simplify data frame
rownames(df) <- df$variable
df <- df[, colnames(df) != "variable"]
```
<br/>

### Format a data frame 
```{r, results="hide"}
# Let's round all values to two decimal digits
x <- lapply(df, function(x) round(as.numeric(x), digits = 2))
x
x <- data.frame(x) # we lost rownames!
rownames(x) <- rownames(df)
df <- x
```
<br/>


### Write a data frame in the file
```{r, results="hide"}
conn <- file("Results.txt", "w")
write.table(df, conn, row.names = T, col.names = NA, sep = "\t")
close(conn)

conn <- file("Results.csv", "w")
write.table(df, conn, row.names = T, col.names = NA, sep = ",")
close(conn)

```
<br/>

### Write a table using as a decimal separator instead of a dot a comma
```{r, results="hide"}
df_comma <- format(df, decimal.mark=",")

conn <- file("Results_comma.txt", "w")
write.table(df_comma, conn, row.names = T, col.names = NA, sep = "\t", quote = F)
close(conn)

```
<br/>



